Simultaneous segmentation and registration of diffusion MR images of the brain driven by active-contours

### 2015-02-18 Publications/2015-NeuroImage/03-introduction.tex
modify font, add lineno

% -*- root: 00-main.tex -*-
\section*{Introduction}
\label{sec:introduction}

Nonlinear registration of images is a challenging task, ubiquitous
  in an endless number of image analysis applications.
The process aims to find a spatial mapping $U$ that aligns the information available
  in two different coordinate systems (generally called
  \emph{reference}, $R$, and \emph{moving}, $M$):
  \begin{align}
  U\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
  \vec{r} &\mapsto \vec{r}' =\vec{r}+u(\vec{r}),
  \label{eq:transform}
  \end{align}
  where $\vec{r}$ denotes a position in the reference domain $R$, $\vec{r}'$ is
  its corresponding location in $M$, and $n$ the dimensionality of images.
Finally, $\vec{u} = u(\vec{r})$ is the displacement of every point with respect
  to the reference domain.
A comprehensive survey by \cite{sotiras_deformable_2013}
  illustrates the wide variety of existing registration methodologies.
They classify the algorithms by the three principal building blocks that can be identified
  among methods: the matching criteria or similarity function to evaluate the proximity of
  the solution, the deformation model that may implement theoretical properties and constraints
  of the mapping in the selected application, and the optimization method or search strategy.
The possibly most widespread matching criteria are those classified as \emph{iconic
  methods} in \citep{sotiras_deformable_2013}.
These algorithms require the definition of an appropriate function on the voxel-wise information
  content of both images.
A second group, called \emph{geometric methods}, align geometrical features derived from
  corresponding objects defined in both spaces, such as landmarks, shapes, surfaces, etc.
Finally, \emph{hybrid methods} map geometrical features and voxel-wise information.

>>>>

% -*- root: 00-main.tex -*-
\section*{Introduction}\label{sec:introduction}
Nonlinear registration of images is a challenging task, ubiquitous
  in an endless number of image analysis applications.
The process aims to find a spatial mapping $U$ that aligns the information available
  in two different coordinate systems (generally called
  \emph{reference}, $R$, and \emph{moving}, $M$):%

  \begin{align}
  U\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
  \vec{r} &\mapsto \vec{r}' =\vec{r}+u(\vec{r}),
  \label{eq:transform}
  \end{align}

  where $\vec{r}$ denotes a position in the reference domain $R$, $\vec{r}'$ is
  its corresponding location in $M$, and $n$ the dimensionality of images.
Finally, $\vec{u} = u(\vec{r})$ is the displacement of every point with respect
  to the reference domain.
A comprehensive survey by \cite{sotiras_deformable_2013}
  illustrates the wide variety of existing registration methodologies.
They classify the algorithms by the three principal building blocks that can be identified
  among methods: the matching criteria or similarity function to evaluate the proximity of
  the solution, the deformation model that may implement theoretical properties and constraints
  of the mapping in the selected application, and the optimization method or search strategy.
The possibly most widespread matching criteria are those classified as \emph{iconic
  methods} in \citep{sotiras_deformable_2013}.
These algorithms require the definition of an appropriate function on the voxel-wise information
  content of both images.
A second group, called \emph{geometric methods}, align geometrical features derived from
  corresponding objects defined in both spaces, such as landmarks, shapes, surfaces, etc.
Finally, \emph{hybrid methods} map geometrical features and voxel-wise information.

### 2015-02-18 Publications/2015-NeuroImage/04-methods.tex
modify font, add lineno

\subsection*{Cost-function derivation}
\label{sec:methods_map}

In a Bayesian framework, the mappings $U$ in \eqref{eq:transform} are
  evaluated based on their posterior probability given the observed data
  $R$.
Using the Bayes' rule, the posterior likelihood is computed as:
  \begin{equation}
  P(U \mid R,\Gamma_{l,m}) = \frac{P(R \mid U,\Gamma_{l,m})\, P(U)}{P(R)},
  \label{eq:bayes_rule}
  \end{equation}
  where $P(R \mid U,\Gamma_{l,m})$ is the data-likelihood, and
  $\Gamma_{l,m}$ are a set of surfaces corresponding to the interfaces
  between piecewise smooth regions $\Omega_i$, such that
  $\Gamma_{l,m} = \partial \Omega_l \cap \partial \Omega_m$ is the
  contour between tissues $l$ and $m$.
Then, the \gls*{map} criterium is applied to find the mapping estimate $\hat{U}$
  that aligns $\Gamma$ into $R$:
  \begin{align}
  \hat{U} &= \underset{U}{\argmax} \left\{ P(U \mid R,\Gamma_{l,m}) \right\} \notag\\
   &= \underset{U}{\argmax} \left\{ P(R \mid U,\Gamma_{l,m})\, P(U) \right\}.
  \label{eq:map_u}
  \end{align}

First, we assume independence between pixels, and thus break down the
  global data likelihood into a product of pixel-wise conditional probabilities:
  \begin{equation}
  P(R \mid U,\Omega) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:bayes_aposteriori}
  \end{equation}
  where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:transform} of the multispectral target
  volume.
For convenience, and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions:
  \begin{equation}
  P\left( \vec{f}' \mid U,\Omega_l \right) = \mathcal{N} \left( \vec{f}' \mid \Theta_l \right),
  \label{eq:likelihood}
  \end{equation}
 	where $\Theta_l = \lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$ are the
 	descriptors of region $\Omega_l$.
We then insert \eqref{eq:likelihood} into \eqref{eq:bayes_rule} to obtain the full
  formulation of the \emph{data term} of our registration model:
 	\begin{align}
  P( R \mid U, \Omega ) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
  &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:pdf}
  \end{align}
  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
  to the descriptors of region $l$ as
  $\mathcal{E}_l(\vec{r}) = \mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.

We choose a Thikonov regularization prior as follows:
  \begin{equation*}
  P(U) = \underset{\vec{r}}{\prod} p(u(\vec{r})) =
  \underset{\vec{r}}{\prod} p_0(u(\vec{r})) \, p_1(u(\vec{r})),
  \end{equation*}
  with further:
  \begin{align*}
  p_0(u(\vec{r})) &= \mathcal{N}( u(\vec{r}) \mid 0, A^{-1}), \notag\\
  p_1(u(\vec{r})) &= \mathcal{N}(  \nabla \cdot u(\vec{r}) \mid 0, B^{-1}),
  \end{align*}
  imposing that the distortion and its gradient have zero
  mean and variance governed by $A$ and $B$.
Since the anisotropy is generally aligned with the imaging axes, these will be simplified
  in the following for the sake of clarity:
  \begin{align}
    p_0(u(\vec{r})) &= \underset{\vec{r}}{\sum} \mathcal{N}( u(\vec{r}) \mid 0,
      (\boldsymbol{\alpha}^{\circ\frac12}\,\vec{I}_n)^{-1}), \notag\\
    p_1(u(\vec{r})) &= \underset{\vec{r}}{\sum} \mathcal{N}( \nabla \cdot u(\vec{r}) \mid 0,
      (\boldsymbol{\beta}^{\circ\frac12}\,\vec{I}_n)^{-1}).
  \label{eq:priors}
  \end{align}

Finally, the \gls{map} problem is adapted into a variational one applying the
  following log-transform:
  \begin{align}
  E(R \mid U) &= -\log \underset{l}{\prod}
  \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( u(\vec{r}))\,p_1( u(\vec{r})) \notag\\
  &= C + \underset{l}{\sum} \int_{\Omega_l}
  \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} \left[ \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2} \right] \,d\vec{r},
  \label{eq:energy}
  \end{align}
  that is the dual expression to the energy functional corresponding
  to a discrete \gls*{acwe} framework \citep{chan_active_2001}
  with anisotropic regularization as studied in
  \citep{nagel_investigation_1986}.
We provide a proof of duality in {\color{red} [REF figshare?]}.

@@

\paragraph*{Deformation model}
\label{sec:deformation_model}
Let us denote $\{\vec{v}_i\}_{i=1 \ldots N_c}$ the vertices of one or several prior
  surface(s).
In our application, these surfaces are triangularized meshes extracted using \emph{FreeSurfer}
  \citep{fischl_freesurfer_2012}.
The transformation from the structural space into the coordinate space of the
  target image is achieved through a dense deformation field $u(\vec{r})$, such that:
  \begin{equation}
  \vec{v}_i' = U\{\vec{v}_i\} = \vec{v}_i + u(\vec{v}_i),
  \label{eq:nodes_tfm}
  \end{equation}
  where $U$ is defined in \eqref{eq:transform}.
Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
  derive $u(\vec{r})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
Densification is achieved through a set of associated basis functions $\psi_k$:
  \begin{equation}
  u(\vec{r}) = \sum_k \psi_k(\vec{r}) \vec{u}_k.
  \label{eq:intp_kernel}
  \end{equation}
%
In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
  of degree 3 ($B_3$).
Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
  $\psi$ by the actual kernel function, the transformation writes:
  \begin{equation}
    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
      B_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
  \label{eq:transformation}
  \end{equation}
  with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.

\paragraph*{Optimization}
\label{sec:gradient_descent}
To find the minimum of the energy functional \eqref{eq:energy},
  we propose a gradient-descent approach with respect to the underlying
  deformation field through the following \gls*{pde}:
  \begin{equation}
  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
  \label{eq:general_gradient_descent}
  \end{equation}
  with $t$ being an artificial time parameter of the contour
  evolution, and $\vec{u}_k$ the parameters supporting the estimate
  $\hat{U}$ of the transformation at the current time point.
Now, we introduce \eqref{eq:energy} in \eqref{eq:general_gradient_descent}:
  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \Big\{
  \int_{\Omega} \underset{l}{\sum} \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} [ \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2} ] \,d\vec{r}
  \Big\}.
  \label{eq:gradient_descent}
  \end{align}

Therefore, we can apply a discretized interpretation of \eqref{eq:shape_gradients}
  to compute the data term in \eqref{eq:gradient_descent} as follows:
  \begin{align}
  \frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \left\{
   \underset{\vec{x} \in \Omega_l}{\sum} \underset{l}{\sum} \mdist{f'}{l} \right\}
  = \underset{i}{\sum}
   \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle,
  \label{eq:gradient_wshape}
  \end{align}
  in this case, the formulation has been adapted to the non-binary case, $\{l,m\}$
  being any pair of neighboring regions, and $\Gamma_{l,m}$ the contour separating
  them such that $\vec{x}' = \vec{v}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$
  and $\vec{n_i}'$ is the unit inward normal to the contour at $c_i'$.

Finally, we can compute:
  \begin{align}
  \frac{\partial \vec{v}_i'}{\partial \vec{u}_k} &= \frac{\partial}{\partial \vec{u}_k}
  \left\{ \vec{v}_i + \sum_k \psi_k(\vec{v}_i) \vec{u}_k \right\}
  = \psi_k(\vec{v}_i)\, \hat{\vec{e}}
  \label{eq:basis_derivative}
  \end{align}
  where $\hat{\vec{e}}$ is the coordinates system's unit vector.
The vertex speeds $\bar{s}_i'$ obtained by computing the shape-gradients \eqref{eq:shape_gradients},
	are then projected to the deformation field in order to obtain the derivatives $\vec{g}_k$
	corresponding to $\vec{u}_k$:
  \begin{equation}
  \vec{g}_{i,k} = \left\langle \frac{\partial}{\partial \vec{u}_k}{\vec{v}_i}', \bar{s}_i'\right\rangle
  = - \left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right] \psi_k(\vec{v}_i)\, \hat{\vec{e}},
  \end{equation}
  then the full gradient evolution equation \eqref{eq:gradient_descent} yields:
  \begin{align}
  \frac{\partial E(\vec{u}_k)}{\partial \vec{u}_k} =
  &\underset{i}{\sum} \vec{g}_{i,k} +2\, \boldsymbol{\alpha} \vec{u}_k
  -2\, \boldsymbol{\beta} \Delta \vec{u}_k,
  \label{eq:gradient_final}
  \end{align}

@@

\subsection*{Experiments and evaluation}
\label{sec:experiments_evaluation}
%
In order to demonstrate the applicability of the registration method, we first
  subjected it to a battery of random tests over synthetic data and then applied
  it on real data.
\autoref{fig:evworkflows} presents the workflow implementing the evaluation instruments.
Besides a visual assessment of the results, we report quantitative evaluations using
  two metrics.
In the case of the phantoms, as we warped data along the three available dimensions,
  we computed the Haussdorf distance, integrating the ``point-to-cell'' method
	implemented by \cite{commandeur_vtk_2011}.
Conversely, the susceptibility-derived distortions only happen along the phase
  encoding axis of the image.
Therefore, a \gls*{swindex} can be computed as the one-to-one distance between corresponding
  vertices of surfaces, weighted by their respective Voronoi area $w_i$:
  \begin{equation}
  sWI = \frac{1}{P} \sum\limits_p^P \frac{1}{A_p} \sum\limits_i^{N_p} w_i\,\|
  \vec{v}_i - \hat{\vec{v}}_i \|
  \end{equation}
  where $\vec{v}_i$ are the locations of the $N_p$ vertices in each $p \in \{1, \dots P\}$
  prior, $A_p$ the total area of surface $p$, and $\hat{\vec{v}}_i$ is the location
  recovered corresponding to the vertex $\vec{v}_i$.

>>>>

\subsection*{Cost-function derivation}\label{sec:methods_map}

In a Bayesian framework, the mappings $U$ in \eqref{eq:transform} are
  evaluated based on their posterior probability given the observed data
  $R$.
Using the Bayes' rule, the posterior likelihood is computed as:

  \begin{equation}
  P(U \mid R,\Gamma_{l,m}) = \frac{P(R \mid U,\Gamma_{l,m})\, P(U)}{P(R)},
  \label{eq:bayes_rule}
  \end{equation}

  where $P(R \mid U,\Gamma_{l,m})$ is the data-likelihood, and
  $\Gamma_{l,m}$ are a set of surfaces corresponding to the interfaces
  between piecewise smooth regions $\Omega_i$, such that
  $\Gamma_{l,m} = \partial \Omega_l \cap \partial \Omega_m$ is the
  contour between tissues $l$ and $m$.
Then, the \gls*{map} criterium is applied to find the mapping estimate $\hat{U}$
  that aligns $\Gamma$ into $R$:

  \begin{align}
  \hat{U} &= \underset{U}{\argmax} \left\{ P(U \mid R,\Gamma_{l,m}) \right\} \notag\\
   &= \underset{U}{\argmax} \left\{ P(R \mid U,\Gamma_{l,m})\, P(U) \right\}.
  \label{eq:map_u}
  \end{align}

First, we assume independence between pixels, and thus break down the
  global data likelihood into a product of pixel-wise conditional probabilities:

  \begin{equation}
  P(R \mid U,\Omega) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:bayes_aposteriori}
  \end{equation}

  where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:transform} of the multispectral target
  volume.
For convenience, and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions:

  \begin{equation}
  P\left( \vec{f}' \mid U,\Omega_l \right) = \mathcal{N} \left( \vec{f}' \mid \Theta_l \right),
  \label{eq:likelihood}
  \end{equation}

 	where $\Theta_l = \lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$ are the
 	descriptors of region $\Omega_l$.
We then insert \eqref{eq:likelihood} into \eqref{eq:bayes_rule} to obtain the full
  formulation of the \emph{data term} of our registration model:

 	\begin{align}
  P( R \mid U, \Omega ) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
  &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:pdf}
  \end{align}

  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
  to the descriptors of region $l$ as
  $\mathcal{E}_l(\vec{r}) = \mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.

We choose a Thikonov regularization prior as follows:

  \begin{equation*}
  P(U) = \underset{\vec{r}}{\prod} p(u(\vec{r})) =
  \underset{\vec{r}}{\prod} p_0(u(\vec{r})) \, p_1(u(\vec{r})),
  \end{equation*}

  with further:

  \begin{align*}
  p_0(u(\vec{r})) &= \mathcal{N}( u(\vec{r}) \mid 0, A^{-1}), \notag\\
  p_1(u(\vec{r})) &= \mathcal{N}(  \nabla \cdot u(\vec{r}) \mid 0, B^{-1}),
  \end{align*}

  imposing that the distortion and its gradient have zero
  mean and variance governed by $A$ and $B$.
Since the anisotropy is generally aligned with the imaging axes, these will be simplified
  in the following for the sake of clarity:

  \begin{align}
    p_0(u(\vec{r})) &= \underset{\vec{r}}{\sum} \mathcal{N}( u(\vec{r}) \mid 0,
      (\boldsymbol{\alpha}^{\circ\frac12}\,\vec{I}_n)^{-1}), \notag\\
    p_1(u(\vec{r})) &= \underset{\vec{r}}{\sum} \mathcal{N}( \nabla \cdot u(\vec{r}) \mid 0,
      (\boldsymbol{\beta}^{\circ\frac12}\,\vec{I}_n)^{-1}).
  \label{eq:priors}
  \end{align}

Finally, the \gls{map} problem is adapted into a variational one applying the
  following log-transform:

  \begin{align}
  E(R \mid U) &= -\log \underset{l}{\prod}
  \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( u(\vec{r}))\,p_1( u(\vec{r})) \notag\\
  &= C + \underset{l}{\sum} \int_{\Omega_l}
  \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} \left[ \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2} \right] \,d\vec{r},
  \label{eq:energy}
  \end{align}

  that is the dual expression to the energy functional corresponding
  to a discrete \gls*{acwe} framework \citep{chan_active_2001}
  with anisotropic regularization as studied in
  \citep{nagel_investigation_1986}.
We provide a proof of duality in {\color{red} [REF figshare?]}.

@@

\paragraph*{Deformation model}\label{sec:deformation_model}
Let us denote $\{\vec{v}_i\}_{i=1 \ldots N_c}$ the vertices of one or several prior
  surface(s).
In our application, these surfaces are triangularized meshes extracted using \emph{FreeSurfer}
  \citep{fischl_freesurfer_2012}.
The transformation from the structural space into the coordinate space of the
  target image is achieved through a dense deformation field $u(\vec{r})$, such that:

  \begin{equation}
  \vec{v}_i' = U\{\vec{v}_i\} = \vec{v}_i + u(\vec{v}_i),
  \label{eq:nodes_tfm}
  \end{equation}

  where $U$ is defined in \eqref{eq:transform}.
Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
  derive $u(\vec{r})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
Densification is achieved through a set of associated basis functions $\psi_k$:

  \begin{equation}
  u(\vec{r}) = \sum_k \psi_k(\vec{r}) \vec{u}_k.
  \label{eq:intp_kernel}
  \end{equation}

In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
  of degree 3 ($B_3$).
Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
  $\psi$ by the actual kernel function, the transformation writes:

  \begin{equation}
    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
      B_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
  \label{eq:transformation}
  \end{equation}

  with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.

\paragraph*{Optimization}
\label{sec:gradient_descent}
To find the minimum of the energy functional \eqref{eq:energy},
  we propose a gradient-descent approach with respect to the underlying
  deformation field through the following \gls*{pde}:

  \begin{equation}
  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
  \label{eq:general_gradient_descent}
  \end{equation}

  with $t$ being an artificial time parameter of the contour
  evolution, and $\vec{u}_k$ the parameters supporting the estimate
  $\hat{U}$ of the transformation at the current time point.
Now, we introduce \eqref{eq:energy} in \eqref{eq:general_gradient_descent}:

  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \Big\{
  \int_{\Omega} \underset{l}{\sum} \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} [ \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2} ] \,d\vec{r}
  \Big\}.
  \label{eq:gradient_descent}
  \end{align}

Therefore, we can apply a discretized interpretation of \eqref{eq:shape_gradients}
  to compute the data term in \eqref{eq:gradient_descent} as follows:

  \begin{align}
  \frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \left\{
   \underset{\vec{x} \in \Omega_l}{\sum} \underset{l}{\sum} \mdist{f'}{l} \right\}
  = \underset{i}{\sum}
   \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle,
  \label{eq:gradient_wshape}
  \end{align}

  in this case, the formulation has been adapted to the non-binary case, $\{l,m\}$
  being any pair of neighboring regions, and $\Gamma_{l,m}$ the contour separating
  them such that $\vec{x}' = \vec{v}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$
  and $\vec{n_i}'$ is the unit inward normal to the contour at $c_i'$.

Finally, we can compute:

  \begin{align}
  \frac{\partial \vec{v}_i'}{\partial \vec{u}_k} &= \frac{\partial}{\partial \vec{u}_k}
  \left\{ \vec{v}_i + \sum_k \psi_k(\vec{v}_i) \vec{u}_k \right\}
  = \psi_k(\vec{v}_i)\, \hat{\vec{e}}
  \label{eq:basis_derivative}
  \end{align}

  where $\hat{\vec{e}}$ is the coordinates system's unit vector.
The vertex speeds $\bar{s}_i'$ obtained by computing the shape-gradients \eqref{eq:shape_gradients},
	are then projected to the deformation field in order to obtain the derivatives $\vec{g}_k$
	corresponding to $\vec{u}_k$:

  \begin{equation}
  \vec{g}_{i,k} = \left\langle \frac{\partial}{\partial \vec{u}_k}{\vec{v}_i}', \bar{s}_i'\right\rangle
  = - \left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right] \psi_k(\vec{v}_i)\, \hat{\vec{e}},
  \end{equation}

  then the full gradient evolution equation \eqref{eq:gradient_descent} yields:

  \begin{align}
  \frac{\partial E(\vec{u}_k)}{\partial \vec{u}_k} =
  &\underset{i}{\sum} \vec{g}_{i,k} +2\, \boldsymbol{\alpha} \vec{u}_k
  -2\, \boldsymbol{\beta} \Delta \vec{u}_k,
  \label{eq:gradient_final}
  \end{align}

@@

\subsection*{Experiments and evaluation}
\label{sec:experiments_evaluation}
%
In order to demonstrate the applicability of the registration method, we first
  subjected it to a battery of random tests over synthetic data and then applied
  it on real data.
\autoref{fig:evworkflows} presents the workflow implementing the evaluation instruments.
Besides a visual assessment of the results, we report quantitative evaluations using
  two metrics.
In the case of the phantoms, as we warped data along the three available dimensions,
  we computed the Haussdorf distance, integrating the ``point-to-cell'' method
	implemented by \cite{commandeur_vtk_2011}.
Conversely, the susceptibility-derived distortions only happen along the phase
  encoding axis of the image.
Therefore, a \gls*{swindex} can be computed as the one-to-one distance between corresponding
  vertices of surfaces, weighted by their respective Voronoi area $w_i$:

  \begin{equation}
  sWI = \frac{1}{P} \sum\limits_p^P \frac{1}{A_p} \sum\limits_i^{N_p} w_i\,\|
  \vec{v}_i - \hat{\vec{v}}_i \|
  \end{equation}

  where $\vec{v}_i$ are the locations of the $N_p$ vertices in each $p \in \{1, \dots P\}$
  prior, $A_p$ the total area of surface $p$, and $\hat{\vec{v}}_i$ is the location
  recovered corresponding to the vertex $\vec{v}_i$.

### 2015-02-18 Publications/2015-NeuroImage/06-discussion.tex
modify font, add lineno

% -*- root: 00-main.tex -*-
\section*{Discussion}
\label{sec:discussion}
We propose a joint registration and segmentation algorithm designed for biomedical
  image analyses that are affected by all or some of the following circumstances:
  i) reference and target images do not present appropriate contrasts to
  	perform volume-based registration reliably, or there is not available an apt image
 		to be used as reference;
  ii) the low resolution of the target volume produces strong partial volume effects
  	that exclude the use of voxel-wise segmentation algorithms and hinder volume-based
  	registration; and,
  iii) the estimation of the unknown warping is explicitly needed and the
    deformation field is smooth enough to be represented with a sufficient number
  	of B-Spline kernels (i.e. the distortion in \gls*{dmri} images,
  	object tracking in time series, etc.).
The registration method performs the segmentation of the target image in homogeneous
  and competing regions defined by explicit active contours.
We chose active contours without edges as we were interested on segmenting homogeneous
  regions of diffusion properties within the brain.
Finally, a B-Spline deformation field supports and keeps track of the evolution of the
  active contours.
Iteratively, the active contours are updated following the instant shape gradients
  \citep{herbulot_segmentation_2006} with a gradient descent strategy.
In section \nameref{sec:methods_background}, we framed our method within the current
  state-of-art and defined the novelty of our approach.

>>>>

% -*- root: 00-main.tex -*-
\section*{Discussion}
\label{sec:discussion}
We propose a joint registration and segmentation algorithm designed for biomedical
  image analyses that are affected by all or some of the following circumstances:
  i) reference and target images do not present appropriate contrasts to
  	perform volume-based registration reliably;
  ii) the low resolution of the target volume produces strong partial volume effects
  	that exclude the use of voxel-wise segmentation algorithms and hinder volume-based
  	registration; and,
  iii) the estimation of the unknown warping is explicitly needed and the
    deformation field is smooth enough to be represented with a sufficient number
  	of B-Spline kernels (i.e. the distortion in \gls*{dmri} images,
  	object tracking in time series, etc.).
The registration method performs the segmentation of the target image in homogeneous
  and competing regions defined by explicit active contours.
We chose active contours without edges as we were interested on segmenting homogeneous
  regions of diffusion properties within the brain.
Finally, a B-Spline deformation field supports and keeps track of the evolution of the
  active contours.
Iteratively, the active contours are updated following the instant shape gradients
  \citep{herbulot_segmentation_2006} with a gradient descent strategy.
In section \nameref{sec:methods_background}, we framed our method within the current
  state-of-art and defined the novelty of our approach.
